As emphasized throughout the thesis, electronic-system designs that ignore
uncertainty are both inefficient and unreliable. Acknowledging this exigent
concern, we have developed a flexible framework for analysis of electronic
systems that are subject to process variation. The framework is capable of
modeling diverse system-level quantities that are dependent on diverse process
parameters, which, in turn, obey diverse probability distributions. Our
methodology delivers a lightweight surrogate for the quantity being analyzed,
thereby providing the designer with a computationally efficient means of
calculating the probability distribution and other characteristics of this
quantity.

Leveraging our general methodology, we have presented a technique aimed at
quantifying transient power and temperature variations of electronic systems
under process uncertainty. The technique has been used to address the
variability in the effective channel length, which is an area of particular
importance. We have drawn a comparison with direct sampling, which has confirmed
the efficiency of our technique in terms of accuracy and speed.

We have also presented a process-variation-aware technique for dynamic
steady-state power and temperature analysis and have proposed a technique for
reliability analysis that seamlessly accounts for the variability in process
parameters and, in particular, for the impact of process variation on the
operating temperature. In this case, we have compared our performance with that
of direct sampling by considering the influence not only of the effective
channel length but also of the gate oxide thickness. The results have shown the
efficiency of our solution with respect to both accuracy and speed.

The low computational demand featured by the proposed framework implies that our
uncertainty analysis can be readily performed inside design-space-exploration
loops, including those targeted at energy and reliability optimization with
temperature-related constraints under process variation. This important virtue
has been demonstrated by considering an energy-driven probabilistic optimization
procedure under temperature- and reliability-related constraints. It has been
shown that temperature has to be treated as a stochastic quantity in order to
make electronic systems reliable.

Finally, note that, even though our framework has been illustrated by
considering the effective channel length, gate oxide thickness, and a number of
specific quantities of interest, it can be readily applied to other problems
that require process variation to be effectively and efficiently taken into
account.
