A bird's eye view of the remaining chapters of the thesis is as follows.

In \cref{background}, we discuss a number of system-level models that are
commonly used in the literature on analysis and design of computer systems and
that are a starting point for the developments presented in this thesis.

In \cref{certainty-development}, we consider techniques for deterministic
system-level analysis of computer systems. These techniques do not take
uncertainty into account; however, they serve as a solid foundation for those
that do. Our attention revolves primarily around power and temperature since
they are of central importance for attaining robustness and energy efficiency.
We develop an accurate and fast approach to dynamic steady-state temperature
analysis of computer systems and apply it in the context of temperature-aware
reliability optimization. Reliability optimization addresses performance
uncertainty by definition; however, it falls in the scope of
\cref{certainty-development} since the accompanying temperature-aware
reliability analysis treats temperature as a deterministic quantity, which is
suboptimal as we discuss in the subsequent chapters.

In \cref{uncertainty-analog-fabrication}, we present our first technique that
addresses uncertainty. Specifically, we develop an approach to analyzing the
variability of process parameters, which is induced by process variation, across
silicon wafers by means of indirect and sparse measurements such as readings
from thermal sensors.

In \cref{uncertainty-analog-development}, we continue working with process
uncertainty and present a technique that is applicable to studying diverse
quantities with respect to process variation. In particular, the approach allows
for progressively analyzing transient power and temperature profiles of computer
systems. Other examples are the dynamic steady-state profile and such critical
metrics as the maximal temperature and energy consumption. They all can be
analyzed from the probabilistic standpoint, and the utility of this virtue is
illustrated by addressing a problem of design-space exploration with
probabilistic constraints related to reliability. Unlike the reliability model
utilized in \cref{certainty-development}, the one presented in
\cref{uncertainty-analog-development} is well aware of process uncertainty and,
therefore, allows for a more adequate treatment of performance uncertainty.

In \cref{uncertainty-digital-development}, we develop another system-level
technique. The tool specializes in tackling digital sources of uncertainty such
as workload variation, which tend to be less regular than the analog ones
including process variation. The proposed technique is exemplified by
quantifying the effect that workload units with unknown processing times have on
the timing-, power-, and temperature-related characteristics of the system under
consideration.

In \cref{uncertainty-digital-management}, we contemplate the future of runtime
management of computer systems under workload variation. In this context, we
perform an early investigation of the utility of advanced prediction techniques
for fine-grained long-range forecasting of the resource usage in large computer
systems.

In \cref{conclusion}, we recapitulate and conclude the thesis.
