Our work has made the following major contributions:

\begin{itemize}

\item
We have proposed a fast and accurate approach to deterministic dynamic
steady-state temperature analysis and have improved deterministic transient
temperature analysis. Leveraging our efficient approach, we have developed a
procedure for reliability optimization targeted at reducing thermal-cycling
fatigue and hence at alleviating aging uncertainty.

\item
We have developed a versatile statistical framework for analyzing the
variability in process parameters across silicon wafers that is induced by the
fabrication process. The key feature of our approach is that it uses indirect
measurements, which streamlines its usage and allows for a reduction in the
costs associated with the production process.

\item
We have developed a versatile probabilistic framework for analyzing system-level
quantities that are affected by process variation. Using our technique, we have
enhanced reliability analysis and optimization of electronic systems by
accounting for process uncertainty in reliability models, which has also
improved the treatment of aging uncertainty.

\item
We have also developed a versatile probabilistic framework for analyzing
system-level quantities affected by workload variation, which tends to engender
a less regular variability than the one that stems from the fabrication process
and thus necessitates a different technique.

\item
We have performed an early yet informative investigation into the utility of the
latest advancements in machine learning for mitigating workload uncertainty at
runtime in the context of resource management. Specifically, we have applied
advanced prediction techniques to fine-grained long-range forecasting of
resource usage in large computer systems.

\end{itemize}

All the proposed solutions have been assessed by extensive experiments, which
have demonstrated the superior performance of our approaches to analysis and
design of electronic systems compared to existing techniques.
