We show how our approach makes it possible to efficiently perform reliability
optimization, based on the thermal cycling (TC) failure mechanism. More exactly,
we propose a temperature-aware task mapping and scheduling technique that
addresses the TC aging effect. Experiments demonstrate the superiority of the
proposed techniques, compared to the state of the art.

The proposed calculation of the SSDTA can be used in a wide range of system
optimizations. One of them is reliability optimization that we discuss in this
section. We perform a temperature-aware task mapping and scheduling in order to
address the thermal cycling fatigue.

\subsection{Platform Model}
\slab{platform-model}

We consider a heterogeneous multicore architecture with a set of processing
elements $\Pi$ defined as the following:
\[
  \Pi = \{ \pi_i = (V_i, f_i, N_{\text{gate} \: i}): \; i = \range{0}{N_p - 1} \}
\]
where $V_i$, $f_i$, and $N_{\text{gate} \: i}$ are the supply voltage,
frequency, and number of gates \cite{liao2005} of the $i$th core, respectively.

\subsection{Application Model}
\slab{application-model}

The periodic application is modeled as a task graph $G = (V, \: E, \: \tau)$
where $V$ is a set of $N_t$ tasks (vertices of the graph), $E$ is a set of data
dependencies between tasks (edges), and $\tau$ is the period of the application,
which we assume to be equal to the deadline. Each pair of a task $v_i \in V$ and
processing element $\pi_j \in \Pi$ is characterized by a tuple $(N_{\text{clock}
\: ij}, C_{\text{eff} \; ij})$, where $N_{\text{clock} \: ij}$ is the number of
clock cycles and $C_{\text{eff} \; ij}$ is the effective switched capacitance.

\subsection{Power Model}
\slab{power-model}

The total power dissipation of a processing element is defined as the sum of the
dynamic and leakage power: $P = P_\text{dyn} + P_\text{leak}$. The dynamic part
is modeled as $P_\text{dyn} = C_\text{eff} f V^2$ where $C_\text{eff}$ is the
effective switched capacitance, $V$ and $f$ are the supply voltage and
frequency, respectively. The leakage part of the power dissipation is defined as
\cite{liao2005}:
\begin{equation} \elab{total-power}
  P_\text{leak}(T) = N_\text{gate} V I_0 \left[ A T^2 e^{\frac{\alpha V + \beta}{T}} + B e^{(\gamma V + \delta)} \right]
\end{equation}
where $T$ and $V$ are the current temperature and supply voltage, respectively,
$N_\text{gate}$ is the number of gates in the circuit, $I_0$ is the average
leakage current at the reference temperature and supply voltage. $A$, $B$,
$\alpha$, $\beta$, $\gamma$, and $\delta$ are the technology-dependent constants
found in \cite{liao2005}.

\subsection{Reliability Model}
\slab{reliability-model}

We address temperature-driven failure mechanisms with the reliability model
presented in \cite{huang2009, xiang2010}. In this paper, our particular focus is
on the thermal cycling (TC) fatigue, which is directly connected to the
temperature variations. The derivation of the model is given in the appendix
(\sref{reliability-optimization}).

Assuming the TC fatigue, the parameters affecting reliability are the amplitude
and number of thermal cycles as well as the maximal temperature. A thermal cycle
is a time interval in which the temperature starts from a certain value and,
after reaching an extremum, returns back.

The mean time to failure (MTTF) of one processing element in the system can be
estimated as the following:
\begin{align} \elab{one-mttf}
  \theta = \frac{\tau}{\sum_{i=0}^{N_m - 1} \frac{1}{N_{c \: i}}}
\end{align}
where $N_m$ is the number of thermal cycles during the application period
$\tau$. $N_{c \: i}$ characterizes the $i$th thermal cycle and is calculated
according to the following expression:
\begin{equation} \elab{cycles-to-failure}
  N_c = A (\Delta T - \Delta T_0)^{-b} e^{\frac{E_a}{k T_\text{max}}}
\end{equation}
where $\Delta T$ is the thermal cycle excursion (the distance between the
minimal and maximal temperatures) and $T_\text{max}$ is the maximal temperature
during the thermal cycle (more details in \sref{reliability-optimization}).

It can be seen that the computation requires the identification of the thermal
cycles with their amplitudes and maximal temperatures. All these are captured by
the SSDTP, which is needed as an input to the reliability optimization.

\subsection{Motivational Example}

Consider an application with six tasks, denoted ``T0''--``T5'', and a
heterogeneous architecture with two cores, labeled ``PE0'' and ``PE1''. The task
graph of the application is given in \fref{task-graph} along with the execution
times for both cores. The period of the application is 0.06 $s$. A first
alternative mapping and schedule, and the resulting SSDTP are shown at the top
of \fref{motivation} (where the height of a task represents its relative dynamic
power consumption). It can be observed that initially PE0 is experiencing three
thermal cycles. If we change the mapping of T5 and move it to PE1, we achieve
two thermal cycles of PE0 instead of three. Finally, if we vary the schedule as
well and change the order of T1 and T3, the number of cycles of PE0 becomes one.
Using the reliability model from \sref{reliability-model}, we observe
improvements in the lifetime of 44.69\% and 54.53\%, respectively, relative to
the initial configuration.

\subsection{Problem Formulation}

The problem formulation is the following:

Given:
\begin{itemize}

\item A multiprocessor system $\Pi$ (\sref{platform-model}).

\item A periodic application $G$ (\sref{application-model}).

\item The floorplan of the chip at the desired level of details, configuration
of the thermal package, and thermal parameters.

\item The parameters of the reliability model (\sref{reliability-model}), i.e.,
the constants $A$, $\Delta T_0$, $b$, $E_a$ (see \elab{cycles-to-failure}).

\end{itemize}

Maximize:
\begin{equation} \elab{fitness-function}
  \mathcal{F} = \min_{i = 0}^{N_p - 1} \theta_i
\end{equation}
such that
\begin{align}
  & t_{\text{end} \: i} \leq \tau, \: \forall i \elab{deadline} \\
  & T_{ij} \leq T_\text{max}, \: \forall i, j \elab{t-max}
\end{align}
where $\theta_i$ is the MTTF of the $i$th processing element given by
\eref{one-mttf}, $t_{\text{end} \: i}$ denotes the end time of the $i$th task,
$\tau$ is the period of the application, and $T_{ij}$ are temperature values in
the SSDTP. \eref{deadline} imposes the application deadline, which we assume to
be equal to the period. \eref{t-max} enforces the constraint on the maximal
temperature in the temperature profile $\mathbb{T} = \{ T_{ij} \}$.

The optimization procedure is based on a genetic algorithm (GA)
\cite{schmitz2004} with the fitness function $\mathcal{F}$ given by
\eref{fitness-function}. The algorithm is outlined in \sref{genetic-algorithm}.

This section contains the derivation of the reliability model discussed in
\sref{reliability-optimization} and the description of the actual optimization
procedure.

\subsubsection{Temperature-Aware Reliability Model}

In our analysis, we use the reliability model presented in \cite{huang2009,
xiang2010}. The model is based on the assumption that the time to failure
$\mathcal{T}$ has a Weibull distribution, i.e., $\mathcal{T} \sim Weibull(\eta,
\beta)$ where $\eta$ and $\beta$ are the scaling and shape parameters,
respectively. The expectation of the distribution is the following:
\begin{equation} \elab{general-mttf}
  \mathbb{E}\left[\mathcal{T}\right] = \eta \; \Gamma(1 + \frac{1}{\beta})
\end{equation}
where $\Gamma$ is the gamma function. $\mathbb{E}\left[\mathcal{T}\right]$ is
the mean time to failure (MTTF) that we denote by $\theta$.

The shape parameter $\beta$ is independent of the temperature variation
\cite{chang2006}, which, however, is not the case with the scaling parameter
$\eta$. Therefore, the distribution varies with the temperature. We can split
the overall period of the application $\tau$ into $N_m$ time intervals $\Delta
t_i$, so that during each time interval $\Delta t_i$ the corresponding $\eta_i$
is a constant:
\begin{equation} \elab{eta-one}
  \eta_i = \frac{\theta_i}{\Gamma(1 + \frac{1}{\beta})}
\end{equation}
where $\theta_i$ is the MTTF in the $i$th time interval as if we had the failure
distribution of this interval all the time. For now the values $\theta_i$ are
unknown and depend on the particular failure mechanism. As it is shown in
\cite{xiang2010}, the reliability function $R(t)$, i.e., the probability of
survival until an arbitrary time $t \geq 0$, can be approximated as the
following:
\[
  R(t) = e^{-(\frac{t}{\tau} \sum_{i=0}^{N_m - 1} \frac{\Delta t_i}{\eta_i})^\beta}
\]
The formula keeps the form of the Weibull distribution with the scaling
parameter equal to:
\begin{equation} \elab{eta-many}
  \eta = \frac{\tau}{\sum_{i=0}^{N_m - 1} \frac{\Delta t_i}{\eta_i}}
\end{equation}
The MTTF with respect to the whole application period can be obtained by
combining \eref{general-mttf}, \eref{eta-one}, and \eref{eta-many}.

As mentioned previously, in order to compute the MTTF, we need to consider the
particular failure mechanism and determine the values $\theta_i$ needed in
\eref{eta-one}. We focus on the thermal cycling fatigue
(\sref{reliability-model}). Assuming this concrete failure model, the duration
$\Delta t_i$, during which the corresponding scaling parameter $\eta_i$ is
constant \eref{eta-one}, is exactly a thermal cycle.

When the system is exposed to identical thermal cycles, the number of such
cycles to failure can be estimated using a modified version of the well-known
Coffin-Manson equation with the Arrhenius term \cite{xiang2010, jedec2010}:
\[
  N_c = A (\Delta T - \Delta T_0)^{-b} e^{\frac{E_a}{k T_\text{max}}}
\]
where $A$ is an empirically determined constant, $\Delta T$ is the thermal cycle
excursion, $\Delta T_0$ is the portion of the temperature range in the elastic
region which does not cause damage, $b$ is the Coffin-Manson exponent, which is
also empirically determined, $E_{a}$ is the activation energy, $k$ is the
Boltzmann constant, and $T_\text{max}$ is the maximal temperature during the
thermal cycle. Over the application period, the system undergoes a number of
different thermal cycles each with its own duration $\Delta t_i$ and each cycle
causes its own damage. Therefore, having $N_m$ thermal cycles characterized by
the number of cycles to failure $N_{c\:i}$ and duration $\Delta t_i$, we can
compute $\theta_i$:
\begin{equation} \elab{mttf-cycle}
  \theta_i = N_{c \: i} \; \Delta t_i
\end{equation}
Taking equations \eref{general-mttf}, \eref{eta-one}, \eref{eta-many}, and
\eref{mttf-cycle} together, we obtain the following expression to estimate the
MTTF of one component in the system:
\begin{align}
  \theta = \frac{\tau}{\sum_{i=0}^{N_m - 1} \frac{1}{N_{c \: i}}}
\end{align}
In order to identify thermal cycles in the temperature curve, we follow the
approach given in \cite{xiang2010} where the rainflow counting method is
employed.

\subsubsection{Optimization Procedure}
\slab{genetic-algorithm}

The optimization procedure is based on a genetic algorithm \cite{schmitz2004}
with the fitness function $\mathcal{F}$ given by \eref{fitness-function}. Each
chromosome is a vector of $2 \times N_t$ elements, where the first half encodes
priorities of the tasks and the second represents a mapping. The population
contains $4 \times N_t$ individuals that are initialized partially randomly and
partially based on the initial temperature-aware solution \cite{xie2006}. In
each generation, a number of individuals, called parents, are chosen for
breeding by the tournament selection with the number of competitors proportional
to the population size. The parents undergo the 2-point crossover with $0.8$
probability and uniform mutation with $0.01$ probability. The evolution
mechanism follows the elitism model where the best individual always survives.
The stopping condition is an absence of improvement within 200 successive
generations.

The fitness of a chromosome, \eref{fitness-function}, is evaluated in a number
of steps. First, the decoded priorities and mapping are given to a list
scheduler that produces schedules for each of the cores. If the application
schedule does not satisfy the deadline, the solution is penalized proportionally
to the delay and is not further evaluated; otherwise, based on the parameters of
the architecture and tasks, a power profile is obtained and the corresponding
SSDTP is computed by our proposed method. If the SSDTP violates the temperature
constraint given by \eref{t-max}, the solution is penalized proportionally to
the amount of violation and not further processed; otherwise, the MTTF of each
core is estimated according to \eref{one-mttf} and the fitness function
$\mathcal{F}$ is computed.

\subsection{Experimental Results}
\slab{reliability-results}

In this section we evaluate the reliability optimization approach described in
\sref{reliability-optimization}, first with a set of synthetic applications and,
finally, using a real-life example.

The experimental setup is the following. Heterogeneous platforms and periodic
applications are generated randomly \cite{dick1998} in such a way that the
execution time of tasks is uniformly distributed between 1 and 10 $ms$ and the
leakage power accounts for 30--60\% of the total power dissipation\footnote{The
parameters of the applications and platforms (task graphs, floorplans, HotSpot
configurations, etc.) used in our experiments are available online at
\cite{liu2011}.}. The linear leakage model is used in the experiments, since, as
discussed in \sref{power-temperature-interdependence}, it provides a good
approximation. The area of one core is 4 $mm^2$, other parameters of the die and
thermal package are given in \tref{parameters}. The temperature constraint
$T_\text{max}$ (see \eref{t-max}) is set to $100^\circ C$. In
\eref{cycles-to-failure} the Coffin-Manson exponent $b$ is set to 6, the
activation energy $E_a$ to 0.5, and the elastic temperature region $\Delta T_0$
to zero \cite{jedec2010}. The coefficient of proportionality $A$ is not
significant, since we are concerned about the relative improvement.

In each of the experiments, we compare the optimized solution with an initial
temperature-aware solution proposed in \cite{xie2006}. This solution consists of
a task mapping and schedule that captures the spatial temperature behavior and
tries to minimize the peak temperature while satisfying the real-time
constraints. The deadline is set to the duration of the initial schedule
extended by 5\%.

In the first set of experiments, we change the number of cores $N_p$ while
keeping the number of tasks $N_t$ per core constant and equal to 20. For each
problem we have generated 20 random task graphs and found the average
improvement of the MTTF over the initial solution ($\scriptstyle
\text{MTTF}_\times$). We also have measured the change in the consumed energy
($\scriptstyle \text{E}_\times$). The results are given in \tref{mttf-cores}
($t$ indicates the optimization time in seconds). It can be seen that the
reliability-aware optimization dramatically increases the MTTF by 13 up to 40
times. Even for large applications with, e.g., 320 tasks deployed onto 16 cores,
a feasible mapping and schedule that significantly improve the lifetime of the
system can be found in an affordable time. Moreover, our optimization does not
impact the energy efficiency of the system.

For the second set of experiments, we keep the quad-core architecture and vary
the size (number of tasks $N_t$) of the application. The number of randomly
generated task graphs per application size is 20. The average improvement of the
MTTF along with the change in the energy consumption are given in
\tref{mttf-tasks}. The observations are similar to those for the previous set of
experiments.

The above experiments have confirmed that our proposed approach is able to
effectively increase the MTTF of the system. The efficiency of this approach is
due to the fast and accurate SSDTP calculation, which is at the heart of the
optimization, and which, due to its speed, allows a huge portion of the design
space to be explored. In order to prove this, we have replaced, inside our
optimization framework, the proposed SSDTP calculation with the calculation
based on HotSpot and based on the SSA, respectively
(\sref{dynamic-steady-state-prior}). The goal is to compare our results with the
results produced using HotSpot and the SSA, after the same optimization time as
needed with the proposed SSDTP calculation technique. The experimental setup is
the same as for the experiments in \tref{mttf-tasks}. The MTTF obtained with
HotSpot and the SSA is evaluated and compared with the MTTF obtained by our
proposed method. The results are summarized in \tref{mttf-comparison}. For
example, the lifetime of the platform running 160 tasks can be extended by more
than 18 times, compared to the initial solution, using our approach, whereas,
the best solutions found with HotSpot and the SSA, using the same optimization
time, are only 2.02 and 5.33 times better, respectively. The reason for the poor
results with HotSpot is the excessively long execution time of the SSDTP
calculation. This allows for a much less thorough investigation of the solution
space than with our proposed technique. In the case of the SSA, the reason is
different. The SSA is fast but also very inaccurate
(\sref{dynamic-steady-state-prior}). The inaccuracy drives the optimization
towards solutions that turn out to be of low quality.

We have seen that our reliability-targeted optimizations have significantly
increased the MTTF without affecting the energy consumption. This is not
surprising, since our optimization will search towards low temperature
solutions, which implicitly means low leakage. In order to further explore this
aspect, we have performed a multi-objective optimization\footnote{The
multi-objective optimization is based on NSGA-II \cite{deb2002}.} along the
dimensions of energy and reliability. An example of the Pareto front averaged
over 20 applications with 80 tasks deployed onto a quad-core platform is given
in \fref{average-pareto}. It can be observed that the variation of energy
is less than 2\%. This means that solutions optimized for the MTTF have an
energy consumption almost identical to those optimized for energy. At the same
time, the difference along the MTTF is huge. This means that ignoring the
reliability aspect one may end up with a significantly decreased MTTF, without
any significant gain in energy.

Finally, we have applied our optimization technique to a real-life example,
namely the MPEG2 video decoder \cite{ffmpeg2011} that is deployed onto a
dual-core platform. The decoder was analyzed and split into 34 tasks. The
parameters of each task were obtained through a system-level simulation using
MPARM \cite{benini2005}. The deadline is set to 40 $ms$ assuming 25 video frames
per second. The solution found with the proposed method improves the lifetime of
the system by 23.59 times with a 5\% energy saving, compared to the initial
solution. The same optimization was solved using HotSpot and the SSA. The best
found solutions are only 5.37 and 11.50 times better than the initial one,
respectively.
