There are a number of related studies that we would like to highlight. Bayesian
inference---which is outlined in \sref{bayesian-statistics}---is utilized in
\cite{zhang2010} for identifying the optimal set of locations on the wafer where
the parameter under consideration should be measured in order to characterize it
with the maximal accuracy. The expectation-maximization algorithm is considered
in \cite{reda2009} in order to estimate missing test measurements. In
\cite{paek2012}, the authors consider an inverse problem focused on the
inference of the power dissipation based on transient temperature maps by means
of Markov random fields. Another temperature-based characterization of power is
developed in \cite{mesa-martinez2007} where a genetic algorithm is employed for
the reconstruction of the power model.

It should be noted that the procedures proposed in \cite{reda2009, zhang2010}
operate on direct measurements, meaning that the output is the same quantity as
the one being measured. In particular, these procedures rely heavily on the
availability of adequate test structures on the dies and are practical only for
secondary quantities affected by process variation such as delays and currents,
but not for the primary ones such as various physical dimensions. Consequently,
they often lead to excessive costs and have a limited range of applications. The
approaches in \cite{mesa-martinez2007, paek2012}, on the other hand,
concentrating on the power dissipation of a single die, are not concerned with
process variation.
