In this section, we elaborate on dynamic steady-state power and temperature
analysis under process variation. Unlike transient analysis, this analysis
cannot be done one time step at a time since the whole repetitive workload needs
to be taken into account at once in order to calculate the corresponding dynamic
steady state. Thus, similar to the contrast between \sref{transient-analysis}
and \sref{dynamic-steady-state-analysis}, the solutions in
\sref{chaos-transient-analysis} and this section differ. However, they still
rely on the same strategy outlined in \sref{chaos-uncertainty-analysis} and
shown in \fref{chaos-overview}.

\subsection{\problemtitle}

The system model is the same as the one in \sref{chaos-transient-analysis}. The
only difference is that the power model is more convenient to be defined as
follows:
\begin{equation} \elab{chaos-power-model-bulk}
  \mp = \f(\vu, \mq).
\end{equation}
The function $\f: \real^\nu \times \real^{\np \times \ns} \to \real^{\np \times
\ns}$ is supposed to return the periodic power profile that corresponds to the
desired periodic workload.

\inputalgorithm{chaos-dynamic-steady-state-iterative}
The solution to probabilistic dynamic steady-state analysis is based on the
deterministic one presented in \sref{dynamic-steady-state-solution}. The power,
temperature, and other vectors that appear in
\sref{dynamic-steady-state-solution} become stochastic in the present context,
which also concerns to the full extent the boundary condition given in
\eref{dynamic-steady-state-condition}. The pseudocode of a procedure that
delivers \mq for a fixed \vu is listed in \aref{dream-dynamic-steady-state}. The
algorithm does not take account of the interdependence between power and
temperature; however, this interdependence can be addressed via one of the
techniques presented in \sref{power-temperature-interplay}. In this section, we
use the iterative approach illustrated in
\aref{dream-dynamic-steady-state-iterative}. For clarity, this last algorithm is
rewritten here as shown in \aref{chaos-dynamic-steady-state-iterative}; the main
difference is that the calculation of power on line~3 is now based on
\eref{chaos-power-model-bulk}.

\begin{remark}
The application of the linear approximation described in
\sref{power-temperature-interplay} is problematic in this case. This technique
is suitable when the only varying parameter is temperature, and all other
parameters have nominal values. In that case, it is relatively easy to decide on
a representative temperature range and apply a curve-fitting procedure. In this
case, however, the power model has multiple parameters stepping far from their
nominal values.
\end{remark}

To recapitulate, in this section, the quantity of interest \g is the dynamic
steady-state temperature profile \mq (or the corresponding power profile \mp, if
preferable). The output of Stage~1 in \fref{chaos-overview} is
\aref{chaos-dynamic-steady-state-iterative}.

\subsection{Surrogate Construction}

At Stage~3 in \fref{chaos-overview}, the procedure delineated in
\sref{chaos-construction} is applied to
\aref{chaos-dynamic-steady-state-iterative} from Stage~1 with respect to the
output of Stage~2,. The algorithm computes \mp or \mq for a given \vu and
returns it in a suitable format.

Following \rref{chaos-multidimensional-output} concerning vector-valued
quantities, \g is treated as an $\np \ns$-element row vector, in which case each
coefficient $\hat{\g}_{\vi}$ in \eref{chaos-expansion} is also such a vector.
The projection matrix in \eref{chaos-projection-matrix} and consequently
\aref{chaos-construction} should be interpreted accordingly: \vg is an $\nq
\times \np \ns$ matrix whose row $j$ is the dynamic steady-state temperature or
power profile computed at point $j$ of the quadrature in
\eref{chaos-coefficient} and reshaped into a row vector. Similarly, $\hat{\vg}$
should be understood as an $\nc \times \np \ns$ matrix whose row $i$ is
coefficient $i$ of the expansion in \eref{chaos-expansion}. Recall that a
certain ordering is assumed to be imposed on quadrature points as well as on
polynomial terms.

The constructed expansion can now be post-processed as required, which is
already the topic of Stage~4 in \fref{chaos-overview}; see
\sref{chaos-processing}
