Consider a generic electronic system that consists of \np heterogeneous
processing elements and is equipped with a thermal package. A processing element
is any active component of the system, that is, any component that consumes
power and dissipates heat. The thermal package is the cooling equipment of the
system, including all passive components, that is, all components that do not
consume power. Processing elements can be identified at different levels of
granularity. For instance, a processing element could be a system on a chip,
a central processing unit, an arithmetic logic unit, cache memory, or a
communication bus. In the majority of the discussions in the thesis, processing
elements correspond to processing units or cores of a multiprocessor system.

The system is characterized by power profiles and temperature profiles, which
are discrete representations of the system's power consumption and heat
dissipation, respectively. A power profile is an $\np \times \ns$ matrix
\begin{equation} \elab{power-profile}
  \mp
  = \left(\range{\vp_1}{\vp_\ns}\right)
  = \left(\range{\vp(t_1)}{\vp(t_\ns)}\right) \in \real^{\np \times \ns}
\end{equation}
that contains \ns samples $\set{\vp_i}_{i = 1}^\ns \subset \real^\np$ of the
power consumption of \np processing elements taken at \ns moments in time
$\set{t_i}_{i = 1}^\ns$ with a sampling interval \dt such that $t_i - t_{i - 1}
= \dt$ for $i = \range{2}{\ns}$. The number \ns is called the number of steps.
Similarly, a temperature profile is an $\np \times \ns$ matrix
\begin{equation} \elab{temperature-profile}
  \mq
  = \left(\range{\vq_1}{\vq_\ns}\right)
  = \left(\range{\vq(t_1)}{\vq(t_\ns)}\right) \in \real^{\np \times \ns}
\end{equation}
that captures the heat dissipation of the processing elements at a number of
equidistant moments in time. Note that, even though the exact time frame and
sampling interval of a power profile or a temperature profile are not included
in the corresponding notation, they are understood from the context.
