Computer systems are omnipresent and omniscient. They penetrate deep into
everyday life and are unsettlingly indispensable and increasingly intelligent at
the tasks entrusted to them. It is then understandable that the design of
computer systems is an acutely difficult and vastly far-reaching process.

One major concern of the designer of a computer system is the presence of
uncertainty due to such phenomena as process, performance, and workload
variation. In many cases, uncertainty is inherent and inevitable, and it leads
to degradation of the quality and efficiency of service in the best case and to
severe faults or burnt silicon in the worst-case scenario. Therefore, it is
crucial to acknowledge and analyze uncertainty and to mitigate its deteriorating
consequences by designing computer systems in such a way that they are well
aware of uncertainty and well equipped to effectively and efficiently tackle it.

We begin by considering techniques for deterministic analysis of certain aspects
of computer systems. These techniques do not take uncertainty into account;
however, they serve as a solid foundation for those that do. Our attention
revolves primarily around power and temperature as they are of central
importance for attaining robustness and energy efficiency. We develop a novel
and fast approach to dynamic steady-state temperature analysis of multiprocessor
systems and apply it in the context of reliability optimization.

We then proceed to developing techniques that address uncertainty. The first
technique of this kind is to infer the variability of process parameters that is
induced by process variation, across silicon wafers based on indirect
measurements such as readings from thermal sensors. The second technique is to
progressively analyze transient power and temperature profiles of multiprocessor
systems with respect to the variability originated from process variation. The
third technique is based on the same machinery as the second but operates from a
different perspective. This technique is suitable for diverse quantities of
interest including dynamic steady-state temperature profiles, and it is
illustrated by addressing a problem of design-space exploration with
probabilistic constraints related to reliability. The fourth technique that we
develop is to efficiently tackle less regular sources of uncertainty than
process variation such as workload variation. This technique is exemplified by
quantifying the effect that workload units with uncertain processing times have
on the timing-, power-, and temperature-related characteristics of
multiprocessor systems.

We finish by focusing on runtime management of computer systems under
uncertainty. In this context, we investigate the utility of advanced prediction
techniques for the purpose of fine-grained long-range forecasting of the
resource usage in cluster clusters. We also develop a computationally efficient
infrastructure for data synthesis in order to facilitate further research in
this direction by making it tractable to experiment with the highly promising
but data-demanding state-of-the-art techniques for modeling and prediction.
