\chapter*{Acknowledgments}
\newcommand\person[2]{#2 #1}

I would like to express my enormous gratitude to Professor \person{Else}{Petru}
and Professor \person{Peng}{Zebo} for their exhaustive and discerning
supervision of my graduate studies at Linköping University. I cannot possibly
imagine a better working environment than the one Petru and Zebo have created
and carefully maintained over all these years in our Embedded Systems
Laboratory.

I am extremely grateful to Professor \person{Marculescu}{Diana} for her caring
and thorough supervision during my four-month visit to Carnegie Mellon
University. Being part of a different research group in a different country was
an invaluable professional experience for me, and Diana made sure that this
experience was also impeccably smooth and indispensably delightful.

I would like to thank Professor \person{Villani}{Mattias} for his insightful
pieces of advice and constant readiness to lend a helping hand of a natural-born
Bayesian to those who occasionally or always find themselves lost in statistics.

I am thankful to \person{Moe}{Anne} for steadily leading our chaotic flock of
graduate students toward doctoral hats, which must be akin in difficulty to
herding cats. I am also thankful to \person{Johansson}{Marie} for her kind help
in every matter.

Over the course of my doctoral education, I have had the pleasure to meet many
other amazing people who have been instrumental in making it possible for me to
get to the finish line of this challenging but gratifying journey.

I thank \person{Rafiliu}{Sergiu} for welcoming me at the train station on my
very first day in Linköping. I thank \person{Bao}{Min} for helping me take my
very first steps as a researcher. I thank \person{Samii}{Soheil} for being an
excellent example of an accomplished senior graduate student whom I aspired to
become myself one day.

I am grateful to \person{SenGupta}{Breeta} for sharing my passion for drawing
and photography. I am grateful to \person{Dutta Bordoloi}{Unmesh} and
\person{Rezine}{Ahmed} for being supportive and reassuring, which I really
needed at times. I am also grateful to \person{Tanasa}{Bogdan} for our engaging
discussions at the whiteboard.

I am thankful to \person{Vasilevskaya}{Maria} for her contagious optimism and
for not letting me forget my Russian. I am thankful to \person{Malmberg}{Lisa}
and \person{Hansson}{Erik} for patiently answering my myriad of questions about
Swedish.

I am particularly grateful to seven people, \person{Horga}{Flavia},
\person{Horga}{Adrian}, \person{Alsaadi}{Sarah}, \person{Lifa}{Adrian},
\person{Holmbäck}{Mikaela}, \person{Maghazeh}{Arian}, and
\person{Arvanitaki}{Antonia}, for our merry gatherings with and without a
reason.

Lastly and most importantly, I cannot thank enough my parents,
\person{Ukhova}{Tatiana} and \person{Ukhov}{Aleksandr}, whose unconditional and
inexhaustible care and support have always been of paramount significance to me.

\vspace{4em}
\hfill
\begin{minipage}{0.35\textwidth}
Ivan Ukhov\\
Linköping, August 2017
\end{minipage}

% \person{Aghaee}{Nima}
% \person{Aminifar}{Amir}
% \person{Dragisic}{Zlatan}
% \person{Ganjei}{Zeinab}
% \person{Ghani Zadegan}{Farrokh}
% \person{Jiang}{Ke}
% \person{Kargén}{Ulf}
% \person{Kärrman}{Åsa}
% \person{Leifler}{Ola}
% \person{Mahfouzi}{Rouhollah}
% \person{Niknafs}{Mina}
% \person{Nikolov}{Dimitar}
% \person{Pantic Dragisic}{Svjetlana}
% \person{Pelayo Danils}{Eva}
% \person{Toczé}{Klervie}
% \person{Vergara}{Ekhiotz}
