\chapter*{Foreword}

I would like to express my enormous gratitude to Professor Zebo Peng and
Professor Petru Else for their exhaustive and discerning supervision of my
graduate studies at Linköping University. I cannot possibly imagine a better
working environment than the one that Zebo and Petru have carefully created and
maintained over all these years in our Embedded Systems Laboratory.

I am extremely grateful to Professor Diana Marculescu for her caring and
thorough supervision during my visit to Carnegie Mellon University. Being part
of a different research group in a different country for several months was an
invaluable professional experience for me, and Diana made sure that this
experience was also impeccably smooth and indispensably delightful.

I would like to thank Professor Mattias Villani for his valuable pieces of
advice and constant readiness to lend a helping hand of a natural-born Bayesian
to those who occasionally find themselves hopelessly lost in statistics.

I am thankful to Anne Moe for attentively shepherding us, graduate students,
toward doctoral hats, which must be akin in difficulty to herding cats. I am
also thankful to Marie Johansson for her kind help with every matter.

Over the course of my doctoral education, I have had the pleasure to meet many
other amazing people who have been instrumental in making it possible for me to
reach the finish line of this challenging and gratifying journey.

I thank Sergiu Rafiliu for welcoming me at the train station on my very first
day in Linköping. I thank Min Bao for helping me take my very first steps as a
researcher. I thank Soheil Samii for being an excellent example of an
accomplished senior graduate student, whom I aspired to become myself one day.

I am grateful to Breeta SenGupta for sharing my passion for drawing and
photography. I am grateful to Unmesh Dutta Bordoloi and Ahmed Rezine for being
supportive and reassuring, which I really needed at times. I am also grateful to
Bogdan Tanasa for our engaging discussions at the whiteboard.

I am thankful to Maria Vasilevskaya for her constantly positive attitude and
contagious optimism. I am also thankful to Lisa Malmberg, Erik Hansson, and Ulf
Kargén for patiently answering my myriad of questions about Swedish.

I am particularly grateful to Flavia Horga, Adrian Horga, Sarah Alsaadi, Adrian
Lifa, Mikaela Holmbäck, Arian Maghazeh, and Antonia Arvanitaki for our numerous
merry gatherings inside and outside of the university's campus.

Lastly and most importantly, I cannot thank enough my parents, Tatiana Ukhova
and Aleksandr Ukhov, whose unconditional and inexhaustible care and support have
always been of paramount significance to me.

\vspace{4em}
\hfill
\noindent
\begin{minipage}{0.35\textwidth}
  Ivan Ukhov\\
  Linköping, October 2017
\end{minipage}
