Due to rapidly increasing power densities, temperature has evolved into an
exigent concern of computer-system designs, and temperature analysis has become
an essential component of design workflows. However, this analysis has
traditionally been a computationally expensive operation. The long computation
times are severely limiting, especially when temperature analysis is to be
performed very many times for such purposes as temperature-aware exploration of
the design space. Therefore, there is a palpable need for developing efficient
techniques for temperature analysis of computer systems. With the above concerns
in mind, we focus on temperature analysis is this chapter. It is important to
realize, however, that temperature analysis is inseparable from power analysis
as power and temperature are interdependent; therefore, power analysis is always
implied whenever we discuss temperature analysis.

There are two types of temperature analysis: \one~transient analysis and
\two~steady-state analysis. The first delivers the temperature that the system
transitions over within an arbitrary time interval staring from an arbitrary
initial condition when it is exposed to an arbitrary, both spatially and
temporally, power distribution. The second delivers the temperature that the
system attains and retains once it has reached a thermal equilibrium under a
power distribution that is spatially arbitrary but temporally constant or
repeating.

Steady-state analysis can be further classified into two categories, which have
already been alluded to: \one~static analysis and \two~dynamic analysis. The
first is concerned with temporally constant power distributions, that is, with
those that do not change over time. In this case, temperature distributions do
not change over time either. The second is concerned with temporary arbitrary
power distributions that are periodic, that is, with those that repeat with
certain periods. In this case, temperature distributions change over time with
the same periods as the corresponding repeating power distributions. The
considered scenario is that the system is exposed to a periodic workload or to
such a workload that can be approximated as periodic. Prominent examples that
have such a periodic behavior are various multimedia applications.

The agenda for this chapter is as follows. In \sref{system-model} and
\sref{power-model}, the system and power models, respectively, are described. In
\sref{temperature-model}, we present the general temperature model that is used
in the above three types of temperature analysis. Transient analysis, static
steady-state analysis, and dynamic steady-state analysis are discussed
separately in \sref{transient-state}, \sref{static-steady-state}, and
\sref{dynamic-steady-state}, respectively. The interdependence between power and
temperature is elaborated on in \sref{power-temperature-interplay}. The
reliability model is presented in \sref{reliability-model}. The importance of
temperature analysis for design-space exploration is covered in
\sref{thermal-cycling-optimization}. \sref{utopia-conclusion} concludes the
chapter.
