The total power consumption of an electrical circuit consists of two main
components: dynamic and static. The dynamic power is due to the actual useful
work done by the circuit. The static power, on the other hand, is due to various
parasitic currents that cannot be entirely eliminated. The total power consumed
by \np processing elements at time $t$ is then written as follows:
\[
  \vp(t) = \vp_\dynamic(t) + \vp_\static(t)
\]
where $\vp_\dynamic \in \real^\np$ and $\vp_\static \in \real^\np$ are the
aforementioned dynamic and static components of the total dissipation of power,
respectively.

The dynamic power is modeled as follows:
\[
  \p_\dynamic = C f V^2
\]
where $C$ is the effective switched capacitance, $f$ is the frequency, and $V$
is the supply voltage of the circuit. The static power is chiefly due to the
leakage current \cite{chandrakasan2000, srivastava2010, juan2011, juan2012},
especially in modern \up{CMOS} transistors. The leakage and, by extension,
static power are modeled as follows \cite{liao2005}:
\begin{equation} \elab{static-power}
  \p_\static = \numberof{g} V I_0 \left(a \q^2 e^{\frac{\alpha V + \beta}{\q}} + b e^{\gamma V + \delta}\right)
\end{equation}
where $\q$ is the current temperature, $V$ is the current supply voltage, $I_0$
is the average leakage current at the reference temperature and reference supply
voltage, and $\numberof{g}$ is the number of gates in the circuit. Quantities
$a$, $b$, $\alpha$, $\beta$, $\gamma$, and $\delta$ are technology-dependent
constants, which can found in \cite{liao2005}.

It can be seen in the above equations that the dynamic power does not depend on
the operating temperature whereas the static power does. The dependence is
highly important; however, we shall ignore it for the moment and return to it in
\sref{power-temperature-interplay} in order to adequately take it into account.
